### 2016-02-08 mcmc.tex
cross-refs within text

% to-do
% -----
% - Kill all issues in this list, below:
% - Deal with all mentions of ``DFM'' or ``HOGG''
% - Go through github issues, if any.
% - Should we make an \acronym{} environment and use it religiously?  I think yes.
% - Do we need to explain the variance on a sampling-based estimator of an integral?  Right now this is just in an end-note (or will be when I write it).  Should it be in the main text?  Probably!
% - mention ESJD
% - mention Gelman--Rubin statistic
% - mention JAGS and STAN in methods section
% - follow up the idea that burn-in doesn't exist at all!  It does for emcee.
% - when we say ``below'' or ``above'' also give a Section Number.
% - need to mention case where there is an MC integral INSIDE the likelihood function
% - need to mention testing the likelihood function
% - would love to have the Note symbols to be links to the notes and the notes to have links back to the main text...!
% - send to friendlies for comments
%   - Gelman
%   - Goodman
%   - Lang
%   - Rix
%   - Bovy
%   - Brewer
% - post on arXiv

@@

In addition to this normalization-insensitive property of \MCMC, in its
simplest forms it can be run without computing any derivatives or
integrals of the function, and (as we will show below) in its simplest
forms it is \emph{extremely easy to implement}.
For all these reasons, \MCMC\ is ideal for sampling posterior pdfs in
the real situations in which scientists find themselves.

@@

All that said, and adhering to the traditions of the \project{Data
  Analysis Recipes} project\note{Every \documentname\ in the \project{Data
    Analysis Recipes} series begins with a rant in which we argue that
  most uses of the methods in question are not appropriate!}, we are
compelled to note at the outset that \MCMC\ is in fact \emph{over-used}.
Because \MCMC\ provably (under assumptions\note{Hint at assumptions
  here.}, some of which will be discussed below) samples the full
posterior pdf in all of parameter space, many investigators use \MCMC\ 
because it will sample \emph{all} of the parameter space
($\pars$-space).
That is, they are using \MCMC\ because they want to
\emph{search the parameter space for good models}.
This is not a good reason to use \MCMC!
Another bad use case is the following:
Because \MCMC\ samples the parameter representatively, it spends most of
its time near very good models; models that (within the confines of
the prior pdf) do a good job of explaining the data.
For this reason, many investigators are using \MCMC\ because it
effectively \emph{optimizes the posterior pdf}, or, for certain
choices of prior pdf, \emph{optimizes the likelihood}.\note{Of course
  a committed Bayesian would argue that any time you are optimizing a
  posterior pdf or optimizing a likelihood, you \emph{should} be
  sampling a posterior pdf.  That is, for some, the fact that when
  someone ``wants'' to optimize, it is actually useful that they
  choose the ``wrong'' tool, because that ``wrong'' tool gives them
  back something far more useful than the output of any optimizer!}
This is another bad reason!

@@

\section{What is a sampling?}

@@

When the sampling $\setof{\pars_k}_{k=1}^K$ is of one of these badly
normalized functions $f(\pars)$---as it usually will be---the
sampling-approximation expectation values are the expectation values
under the properly-normalized corresponding pdf, even though you might
\emph{never learn that normalization}.
When we run \MCMC\ sampling on $f(\pars)$, a function that differs from a
pdf $p(\pars)$ by some unknown normalization constant, then the sampling
permits computation of the following kinds of quantities:
\begin{eqnarray}
E[q] &\equiv& \frac{\int q(\pars)\,f(\pars)\,\dd \pars}{\int f(\pars)\,\dd \pars}
\\
E[q] &\approx& \frac{1}{K}\,\sum_{k=1}^K q(\pars_k)
\label{eq:expectationvalue}
\quad .
\end{eqnarray}
That is, the sampling can be constructed (as we will show below) from
$f(\pars)$ directly, and it permits you to compute expectation values
without ever requiring you to integrate either the numerator integral
or the denominator integral, both of which are generally
intractable.\note{As we will
  see, the intractability comes from various directions, but one is
  that the dimension of $\pars$ gets large in most realistic situations.
  Another is that the support of $p(\pars)$ tends to be, in normal
  inference situations, \emph{far smaller} than the full domain of
  $\pars$.
  That is, the pdf is at or very close to zero over all but some tiny
  and very hard-to-find part of the space.}

The ``correctness'' of a sampling is defined (above) in terms of its use in
  performing integrals or approximate computation of integrals.
In a deep sense, the \emph{only} thing a sampling is good for is
  computing integrals.
There are many uses of the \MCMC\ sampling, some of them good and some of them bad.
Most of the good or sensible uses will somehow involve integration.

% in the next paragraph I adopt $D$ to be the dimensionality of the space

@@

Finally---and perhaps most importantly---a magical property of a sampling
  (in a $D$-dimensional space)
  is that if some of your $D$ dimensions are totally uninteresting
  (nuisance parameters, if you will)
  and some of your $D$ dimensions are of great interest,
  the sampling in the full $D$-space is trivially converted into a sampling in the subspace of interest:
You just drop from each $\pars_k$ vector (or blob or list) the dimensions of no interest!
That is, the projection of the sampling to the subspace of interest produces
  a sampling of the marginalized pdf, marginalizing (or projecting) out the nuisance parameters.%
  \note{This point about marginalization is, once again, a use of the sampling to perform an \emph{integral};
        the marginalized pdf is obtained from the full pdf by an integration.
        The sampling performs this integration automatically.}
That is extremely important for inference,
  where there are always parameters with very different levels of importance to the scientific conclusions.
This point will return again below in \sectionname~\ref{sec:results}.

Although the discussion in this \documentname\ is general, the most
common use of \MCMC\ sampling (for us, anyway) is in probabilistic
inference.
For this reason, we will often refer to the function $f(\pars)$
colloquially as ``the posterior pdf''\note{We will also sometimes
  refer to expectations under $f(\pars)$ as posterior means, and
  medians as median-of-posterior values, and so on.} even though it is
implicitly ill-normalized and might not be a posterior pdf in any
sense.
We will also occasionally assume---just because it is true in
inference---that the function $f(\pars)$ is the product of two
functions, one called ``the prior pdf'' and one called ``the
likelihood''.
Again, this usage is colloquial and is only strictly correct in
inference contexts with proper inputs.
That the function $f(\pars)$ can be thought of as a product of a
prior pdf and a likelihood is only necessary for what follows in
the context of advanced sampling techniques like tempering or
nested sampling, both mentioned briefly below.

@@

\section{Metropolis--Hastings MCMC}

@@

This algorithm---and indeed any \MCMC\ algorithm---produces a
  \emph{biased random walk} through parameter space.
It is a random walk for the same reason that it is ``Markov'':
The step it makes to position $\pars_{k+1}$ depends only on the state
  of the sampler at position $\pars_k$ (and no previous state).
It is a biased random walk, biased by the acceptance algorithm
  involving the ratios of function values; this acceptance rule biases
  the random walk such that the amount of time spent in the neighborhood
  of location $\pars$ is proportional to $f(\pars)$.
Because of this local (Markov) property, the nearby samples are not
  independent; the algorithm only produces fair samples in the limit of
  arbitrary run time, and two samples are only independent when they are
  sufficiently separated in the chain (more on this below).

The principal user-settable knob in \MH\ \MCMC\ is the proposal
  pdf $q(\pars'\given\pars)$.
A typical choice is a multi-variate Gaussian distribution for $\pars'$
  centered on $\pars$ with some simple (diagonal, perhaps) variance
  tensor.
We will discuss the choice and tuning of this proposal distribution
  below.

@@

In what follows, when we discuss tuning of the proposal pdf, we will
  often cycle through the parameter blob components and propose changes
  in only one dimension or element or component at a time.
That is, at step $k+1$ you might only propose a one-dimensional move
  in the $i$th dimension, not in all $D$ dimensions of the parameter
  space.
That won't change anything significant; but if you are implementing
  this right now you might want to do that.
It will help us tune the proposal pdf (\sectionname~\ref{sec:tuning})
and diagnose problems (\sectionname~\ref{sec:troubleshooting}).

One note to make here
  is that you must \emph{propose}
  in the same coordinates (parameterization or transformation of parameters)
  as that in which your priors are \emph{specified},
  or else multiply in a (possibly nasty) Jacobian.
That is, if your prior is ``flat in $\pars$''
  but you find it easier to run your sampler by proposing steps in $\ln \pars$,
  if you don't modify your acceptance fraction by the Jacobian,
  your \emph{real} prior won't be flat in $\pars$.
These think-os can get subtle;
  the best way to avoid them
  is to write your code and your likelihood function
  and your prior pdf function and your proposal distribution
  all in precisely the same parameterization.
We will return to this point below in our comments on functional testing.

As we discuss briefly below in \sectionname~\ref{sec:methods},
  there are many \MCMC\ methods more advanced than M--H.
However, they all share characteristics with M--H
  (initialization, tuning, judging convergence),
  such that it is very valuable to understand \MH\ well before using anything more advanced.
Furthermore, a scientist new to \MCMC\ benefits enormously from
  building, tuning, and using her or his own \MCMC\ software.
One piece of advice we give then,
  to the new user of \MCMC,
  is to code up, tune, and use a \MH\ \MCMC\ sampler for a scientific project.
A huge amount is learned in doing this,
  and it is very often the case that the home-built \MH\ \MCMC\ does everything needed;
  you often don't need any more advanced tool.
Only after you have concluded that your home-built \MH\ \MCMC\ sampler
  is not suited to your project
  (or not developed properly into a properly versatile software package)
  should you download and start to use any professionally developed alternative.
That is, even if---in the end---%
  you want to leave the sampling code to the experts and run an industrial-strength code,
  it is still valuable to build your own,
  given the simplicity of the algorithm,
  and given the intuition you gain by doing it (at least once) yourself.\note{%
    We also make this point in a previous \documentname\ (\citealt{fittingaline};
    the ambitious reader will find in that \documentname\ a solid example project
    for using \MCMC\ in a real context.}

@@

A sampler with a smaller autocorrelation time is just better; you have
to do fewer $f(\pars)$ calls per independent sample, and you have to
run less time to get accurate sampling-based integral estimates.
A sampler that takes an independent sample every time would have an
autocorrelation time of unity, which is the best possible value; this
optimal sampling is only possible for problems where the sampling is
analytic (for example if the posterior is perfectly Gaussian with
known mean and variance).
In general, the best sampler will be different for different problems,
and we will (below) tune the samplers we have to do the best on the
problems we have; this tuning will also be problem-specific.
There are many heuristic bases on which different samplers might be
compared or tuned, but fundamentally it is lower autocorrelation time
that separates good samplers from bad ones and is the ultimate basis
on which we compare performance.\note{HOGG: Bash Dunkley and anyone
  else who has used something else to compare.}

@@

In a deep sense, the only scalar that makes sense to optimize,
  when choosing proposal distribution (or, loosely speaking, step size),
  is the autocorrelation time, described above.
The optimal step size is the step size that makes for the shortest autocorrelation time.
This is easy to state, but hard to use in practice;
  being a second-order statistic of the \MCMC\ chain,
  the autocorrelation time is a hard thing to measure without a lot of data,
  so it is hard to quickly estimate it and adjust,
  much less put it into some kind of optimization loop.
We usually use proxies for this of various kinds.

The simplest heuristic proxy statistic for tuning is the \emph{acceptance fraction}.
If you are accepting almost all proposed steps, your step sizes are too small on average.
If you are accepting almost none, your step sizes are too large.
The Goldilocks value is between a half and about a quarter,
  with an argument floating around that it should be 0.234 for best
  performance in high dimensional problems\note{%
    The argument and assumptions underlying the 0.234 fraction
    (and higher acceptance fractions at lower numbers of dimensions)
    are laid out in \citet{gelman234}.}
  (though you could never tune it precisely enough to warrant that third digit of accuracy).
In the burn-in phase (discussed below) of an \MCMC\ run,
  it makes sense to track the acceptance ratio,
  and adjust the proposal distribution variance as you get acceptance ratios
  that are far from the Goldilocks ratio.
This process can be automated easily;
  such automation is part of many projects that use \MCMC.

@@

\section{Initialization and burn-in}

@@

One thing we can keep in mind to guide us in this is what we said above,
  which is that samplings are good for doing \emph{integrals}.
We should endeavor to use as our ``results'' outputs from the \MCMC\ 
  that are based on integrals computed with the sampling.
This includes expectations, medians, quantiles, one-dimensional histograms,
  and multi-dimensional histograms.
This does \emph{not} include the ``best'' sample or a mode or optimum.
The latter things are not necessarily illegitimate outputs of the \MCMC,
  but they do not make best use of the fact that the sampling is a tool for integration,
  and we do not recommend them.

@@

As we mentioned above, scientists often like to report the ``best sample'';
  that is, the sample with the highest posterior pdf value.
This is not usually a good idea.
For one, \MCMC\ is a sampling algorithm, not an optimization algorithm;
  there are no guarantees that it will find the optimum of the posterior pdf in reasonable time.%
  \note{The sampler samples the pdf fairly.
    The tallest peak in the posterior is not guaranteed to be---%
    and in general is not---%
    also the peak with the greatest posterior ``mass''.
    That is, the samples in a fair sampling will not necessarily come
    predominantly from the \emph{tallest} peak in the posterior pdf.
    They will come from the peak with the greatest total integral
    of the likelihood over the prior.
    These issues might sound ``academic'',
    but when the number of parameters gets large (greater than, say, 10),
    many normal intuitions one might have (if any) about what is reasonably likely in a pdf
    are regularly violated.}
For two, the closeness of the best sample to the posterior mode is a strong function of sample size,
  and with very bad scaling.%
  \note{If you have a $K$-point sampling in a $D$-dimensional parameter space,
    there is some best sample $\pars_k^{\best}$.  Now imagine that you want to make
    that best sample 10 times better---that is, you want to find a point $\pars^{\better}$
    that is ten times closer to the true optimum.  How many more samples do you need to take?
    In the limit that $K$ is large, on average, you have to take a factor of
    $10^D$ more samples!
    Any reasonable optimizer is far, far better than this (in terms of compute time).}
For three, if you just want an optimum, use an optimizer!
That is, we \emph{strongly advise against} reporting,
  as ``the'' measurement or ``the result'',
  the parameter value or values for the best sample.
Giving some posterior samples is a good idea (see below).
If you feel drawn to give the best sample,
  instead give the parameter values found by optimizing the posterior PDF,
  starting at the best sample as an initialization
  (and sacrifice your probabilism\note{something}).
This optimal-posterior parameter value is called the
  ``maximum \foreign{a posteriori}'' or ``MAP'' value for the parameters;
  it is like a maximum-likelihood value but regularized by the prior.
It is an estimator with some good (and some bad) properties, but it is not the point of \MCMC,
  and therefore outside the scope of this \documentname.

@@

In principle the only output from the sampling that safely gives both
probabilistic information about the result of the inference and also
good-fitting models is a few---randomly chosen---example samples.
We strongly recommend this\note{HOGG:  And we do this WHERE? Citation!};
in particular this is a better thing to do
than to give the ``best sample'', which isn't even guaranteed to be
near the bulk of the posterior pdf or the samples therefrom.

@@

As we have discussed above, \MCMC\ methods are proven to be correct ``in the limit''; that is, when
an infinite number of samples have been taken.
When the function $f(\pars)$ has challenging properties, the results
can approach this limit very slowly (or not at all\note{HOGG: discuss
  situations in which \MCMC\ will not work---related to support.}).
For this reason, in the finite duration of a scientific project, most
users of \MCMC\ have one of two (related) problems:
Either \textsl{(1)}~it is taking too long---it is taking too many steps or
executing too many calls of the function $f(\pars)$ to get independent
samples---or \textsl{(2)}~it is not
exploring the full parameter space---there are local optima (or bad
local geometry around those optima) ``trapping'' the algorithm.

@@

\paragraph{Ensembles:}
As we discuss more below, one of the critical challenges in the use of
\MCMC\ is tuning the proposal pdf $q(\pars'\given\pars)$.
In problems in which all of the parameters (all of the components or entries of $\pars$) are somehow ``equivalent'',
or the parameters can be seen as components of a vector in a vector
space, there are \emph{ensemble methods} that make use of not just one
random-walker but instead many walkers to automatically generate a properly tuned
proposal distribution.
These methods make use of the distribution of a set of independent
walkers in the parameter space ($\pars$-space) to gauge the typical step-sizes and
directions at which new proposals should be made, obviating or greatly
reducing the tuning requirements.
We are co-authors on one of these methods, \project{emcee}\note{HOGG:
  ref emcee}, which is popular in the astrophysical sciences.
Because ensemble methods are updating independent \MCMC\ chains, they
are often easy to parallelize to make use of multiple cores.

@@

A trivial kind of Gibbs sampling was mentioned above in the
discussion of \MH\ \MCMC:
Even when running vanilla \MH\ \MCMC, it is often useful to update only
one parameter at a time; that is, do only axis-aligned moves in the
$\pars$-space, cycling through axes as you go from step to step.
This helps with tuning and diagnosis, but it also permits a good
implementation to capitalize on simpler calculations for the simpler
parameter directions.\note{%
  The point is that sometimes there are some parameters that only
  affect a very small number of data points, or parameters that change
  the model or the likelihood function and prior values in a very
  simple-to-compute way.  If such simple parameters exist, it makes
  sense to write your likelihood and prior code to capitalize on these
  simplicities.  This is not always trivial, because it may involve
  caching parts of the calculation and so on.  Details would be beyond
  our scope.}

@@

% HOGG say:  I have no idea what the comment here (below) means, even though I wrote it!
% ---
% Give a sense of the difficulties you get into if you fail here; like
% discussion with Sven Kreiss on 2013-02-14...
% ---

@@

\section{Troubleshooting and advice}\label{sec:troubleshooting}

>>>>

% to-do
% -----
% - Kill all issues in this list, below:
% - Deal with all mentions of ``DFM'' or ``HOGG''
% - Go through github issues, if any.
% - Systematically cite chapters of Brooks \etal in each section.
% - mention ESJD
% - mention Gelman--Rubin statistic
% - mention JAGS and STAN in methods section
% - follow up the idea that burn-in doesn't exist at all!  It does for emcee.
% - need to mention case where there is an MC integral INSIDE the likelihood function
% - need to mention testing the likelihood function, by, say, 1-d scans back and forth.
% - would love to have the Note symbols to be links to the notes and the notes to have links back to the main text...!
% - send to friendlies for comments
%   - Gelman
%   - Goodman
%   - Lang
%   - Rix
%   - Bovy
%   - Brewer
% - post on arXiv

@@

In addition to this normalization-insensitive property of \MCMC, in its
simplest forms it can be run without computing any derivatives or
integrals of the function, and (as we will show below in \sectionname~\ref{sec:MH}) in its simplest
forms it is \emph{extremely easy to implement}.
For all these reasons, \MCMC\ is ideal for sampling posterior pdfs in
the real situations in which scientists find themselves.

@@

All that said, and adhering to the traditions of the \project{Data
  Analysis Recipes} project\note{Every \documentname\ in the \project{Data
    Analysis Recipes} series begins with a rant in which we argue that
  most uses of the methods in question are not appropriate!}, we are
compelled to note at the outset that \MCMC\ is in fact \emph{over-used}.
Because \MCMC\ provably (under assumptions\note{HOGG: Hint at assumptions
  here.}, some of which will be discussed) samples the full
posterior pdf in all of parameter space, many investigators use \MCMC\ 
because it will sample \emph{all} of the parameter space
($\pars$-space).
That is, they are using \MCMC\ because they want to
\emph{search the parameter space for good models}.
This is not a good reason to use \MCMC!
Another bad use case is the following:
Because \MCMC\ samples the parameter representatively, it spends most of
its time near very good models; models that (within the confines of
the prior pdf) do a good job of explaining the data.
For this reason, many investigators are using \MCMC\ because it
effectively \emph{optimizes the posterior pdf}, or, for certain
choices of prior pdf, \emph{optimizes the likelihood}.\note{Of course
  a committed Bayesian would argue that any time you are optimizing a
  posterior pdf or optimizing a likelihood, you \emph{should} be
  sampling a posterior pdf.  That is, for some, the fact that when
  someone ``wants'' to optimize, it is actually useful that they
  choose the ``wrong'' tool, because that ``wrong'' tool gives them
  back something far more useful than the output of any optimizer!}
This is another bad reason!

@@

\section{What is a sampling?}\label{sec:sampling}

@@

When the sampling $\setof{\pars_k}_{k=1}^K$ is of one of these badly
normalized functions $f(\pars)$---as it usually will be---the
sampling-approximation expectation values are the expectation values
under the properly-normalized corresponding pdf, even though you might
\emph{never learn that normalization}.
When we run \MCMC\ sampling on $f(\pars)$, a function that differs from a
pdf $p(\pars)$ by some unknown normalization constant, then the sampling
permits computation of the following kinds of quantities:
\begin{eqnarray}
E[q] &\equiv& \frac{\int q(\pars)\,f(\pars)\,\dd \pars}{\int f(\pars)\,\dd \pars}
\\
E[q] &\approx& \frac{1}{K}\,\sum_{k=1}^K q(\pars_k)
\label{eq:expectationvalue}
\quad .
\end{eqnarray}
That is, the sampling can be constructed (as we will show below in \sectionname~\ref{sec:MH}) from
evaluations of $f(\pars)$ directly, and it permits you to compute expectation values
without ever requiring you to integrate either the numerator integral
or the denominator integral, both of which are generally
intractable.\note{As we will
  see, the intractability comes from various directions, but one is
  that the dimension of $\pars$ gets large in most realistic situations.
  Another is that the support of $p(\pars)$ tends to be, in normal
  inference situations, \emph{far smaller} than the full domain of
  $\pars$.
  That is, the pdf is at or very close to zero over all but some tiny
  and very hard-to-find part of the space.}

The ``correctness'' of a sampling is defined (above in \sectionname~\ref{sec:sampling}) in terms of its use in
  performing integrals or approximate computation of integrals.
In a deep sense, the \emph{only} thing a sampling is good for is
  computing integrals.
There are many uses of the \MCMC\ sampling, some of them good and some of them bad.
Most of the good or sensible uses will somehow involve integration.

@@

Finally---and perhaps most importantly---a magical property of a sampling
  (in a $D$-dimensional space)
  is that if some of your $D$ dimensions are totally uninteresting
  (nuisance parameters, if you will)
  and some of your $D$ dimensions are of great interest,
  the sampling in the full $D$-space is trivially converted into a sampling in the subspace of interest:
You just drop from each $\pars_k$ vector (or blob or list) the dimensions of no interest!
That is, the projection of the sampling to the subspace of interest produces
  a sampling of the marginalized pdf, marginalizing (or projecting) out the nuisance parameters.%
  \note{This point about marginalization is, once again, a use of the sampling to perform an \emph{integral};
        the marginalized pdf is obtained from the full pdf by an integration.
        The sampling performs this integration automatically.}
That is extremely important for inference,
  where there are always parameters with very different levels of importance to the scientific conclusions.
This point will return again below (\sectionname~\ref{sec:results}).

Although the discussion in this \documentname\ is general, the most
common use of \MCMC\ sampling (for us, anyway) is in probabilistic
inference.
For this reason, we will often refer to the function $f(\pars)$
colloquially as ``the posterior pdf''\note{We will also sometimes
  refer to expectations under $f(\pars)$ as posterior means, and
  medians as median-of-posterior values, and so on.} even though it is
implicitly ill-normalized and might not be a posterior pdf in any
sense.
We will also occasionally assume---just because it is true in
inference---that the function $f(\pars)$ is the product of two
functions, one called ``the prior pdf'' and one called ``the
likelihood''.
Again, this usage is colloquial and is only strictly correct in
inference contexts with proper inputs.
That the function $f(\pars)$ can be thought of as a product of a
prior pdf and a likelihood is only necessary for what follows in
the context of advanced sampling techniques like tempering or
nested sampling, both mentioned briefly below (\sectionname~\ref{sec:methods}).

@@

\section{Metropolis--Hastings MCMC}\label{sec:MH}

@@

This algorithm---and indeed any \MCMC\ algorithm---produces a
  \emph{biased random walk} through parameter space.
It is a random walk for the same reason that it is ``Markov'':
The step it makes to position $\pars_{k+1}$ depends only on the state
  of the sampler at position $\pars_k$ (and no previous state).
It is a biased random walk, biased by the acceptance algorithm
  involving the ratios of function values; this acceptance rule biases
  the random walk such that the amount of time spent in the neighborhood
  of location $\pars$ is proportional to $f(\pars)$.
Because of this local (Markov) property, the nearby samples are not
  independent; the algorithm only produces fair samples in the limit of
  arbitrary run time, and two samples are only independent when they are
  sufficiently separated in the chain (more on this below in \sectionname~\ref{sec:convergence}).

The principal user-settable knob in \MH\ \MCMC\ is the proposal
  pdf $q(\pars'\given\pars)$.
A typical choice is a multi-variate Gaussian distribution for $\pars'$
  centered on $\pars$ with some simple (diagonal, perhaps) variance
  tensor.
We will discuss the choice and tuning of this proposal distribution
  below (\sectionname~\ref{sec:tuning} and \sectionname~\ref{sec:trouble}).

@@

In what follows, when we discuss tuning of the proposal pdf, we will
  often cycle through the parameter blob components and propose changes
  in only one dimension or element or component at a time.
That is, at step $k+1$ you might only propose a one-dimensional move
  in the $i$th dimension, not in all $D$ dimensions of the parameter
  space.
That won't change anything significant; but if you are implementing
  this right now you might want to do that.
It will help us tune the proposal pdf (\sectionname~\ref{sec:tuning})
and diagnose problems (\sectionname~\ref{sec:trouble}).

One note to make here
  is that you must \emph{propose}
  in the same coordinates (parameterization or transformation of parameters)
  as that in which your priors are \emph{specified},
  or else multiply in a (possibly nasty) Jacobian.
That is, if your prior is ``flat in $\pars$''
  but you find it easier to run your sampler by proposing steps in $\ln \pars$,
  if you don't modify your acceptance fraction by the Jacobian,
  your \emph{real} prior won't be flat in $\pars$.
These think-os can get subtle;
  the best way to avoid them
  is to write your code and your likelihood function
  and your prior pdf function and your proposal distribution
  all in precisely the same parameterization.
We will return to this point below in our comments on testing (\sectionname~\ref{sec:trouble}).

As we discuss briefly below (\sectionname~\ref{sec:methods}),
  there are many \MCMC\ methods more advanced than M--H.
However, they all share characteristics with M--H
  (initialization, tuning, judging convergence),
  such that it is very valuable to understand \MH\ well before using anything more advanced.
Furthermore, a scientist new to \MCMC\ benefits enormously from
  building, tuning, and using her or his own \MCMC\ software.
One piece of advice we give then,
  to the new user of \MCMC,
  is to code up, tune, and use a \MH\ \MCMC\ sampler for a scientific project.
A huge amount is learned in doing this,
  and it is very often the case that the home-built \MH\ \MCMC\ does everything needed;
  you often don't need any more advanced tool.
Only after you have concluded that your home-built \MH\ \MCMC\ sampler
  is not suited to your project
  (or not developed properly into a properly versatile software package)
  should you download and start to use any professionally developed alternative.
That is, even if---in the end---%
  you want to leave the sampling code to the experts and run an industrial-strength code,
  it is still valuable to build your own,
  given the simplicity of the algorithm,
  and given the intuition you gain by doing it (at least once) yourself.\note{%
    We also make this point in a previous \documentname\ (\citealt{fittingaline};
    the ambitious reader will find in that \documentname\ a solid example project
    for using \MCMC\ in a real context.}

@@

A sampler with a smaller autocorrelation time is just better; you have
to do fewer $f(\pars)$ calls per independent sample, and you have to
run less time to get accurate sampling-based integral estimates.
A sampler that takes an independent sample every time would have an
autocorrelation time of unity, which is the best possible value; this
optimal sampling is only possible for problems where the sampling is
analytic (for example if the posterior is perfectly Gaussian with
known mean and variance).
In general, the best sampler will be different for different problems,
and we will (below; \sectionname~\ref{sec:methods}) tune the samplers we have to do the best on the
problems we have; this tuning will also be problem-specific.
There are many heuristic bases on which different samplers might be
compared or tuned, but fundamentally it is lower autocorrelation time
that separates good samplers from bad ones and is the ultimate basis
on which we compare performance.\note{HOGG: Bash Dunkley and anyone
  else who has used something else to compare.}

@@

In a deep sense, the only scalar that makes sense to optimize,
  when choosing proposal distribution (or, loosely speaking, step size),
  is the autocorrelation time, described above (\sectionname~\ref{sec:convergence}).
The optimal step size is the step size that makes for the shortest autocorrelation time.
This is easy to state, but hard to use in practice;
  being a second-order statistic of the \MCMC\ chain,
  the autocorrelation time is a hard thing to measure without a lot of data,
  so it is hard to quickly estimate it and adjust,
  much less put it into some kind of optimization loop.
We usually use proxies for this of various kinds.

The simplest heuristic proxy statistic for tuning is the \emph{acceptance fraction}.
If you are accepting almost all proposed steps, your step sizes are too small on average.
If you are accepting almost none, your step sizes are too large.
The Goldilocks value is between a half and about a quarter,
  with an argument floating around that it should be 0.234 for best
  performance in high dimensional problems\note{%
    The argument and assumptions underlying the 0.234 fraction
    (and higher acceptance fractions at lower numbers of dimensions)
    are laid out in \citet{gelman234}.}
  (though you could never tune it precisely enough to warrant that third digit of accuracy).
In the burn-in phase (discussed below in \sectionname~\ref{sec:initialization}) of an \MCMC\ run,
  it makes sense to track the acceptance ratio,
  and adjust the proposal distribution variance as you get acceptance ratios
  that are far from the Goldilocks ratio.
This process can be automated easily;
  such automation is part of many projects that use \MCMC.

@@

\section{Initialization and burn-in}\label{sec:initialization}

@@

One thing we can keep in mind to guide us in this is what we said above (\sectionname{sec:sampling}),
  which is that samplings are good for doing \emph{integrals}.
We should endeavor to use as our ``results'' outputs from the \MCMC\ 
  that are based on integrals computed with the sampling.
This includes expectations, medians, quantiles, one-dimensional histograms,
  and multi-dimensional histograms.
This does \emph{not} include the ``best'' sample or a mode or optimum.
The latter things are not necessarily illegitimate outputs of the \MCMC,
  but they do not make best use of the fact that the sampling is a tool for integration,
  and we do not recommend them.

@@

As we mentioned above, scientists often like to report the ``best sample'';
  that is, the sample with the highest posterior pdf value.
This is not usually a good idea.
For one, \MCMC\ is a sampling algorithm, not an optimization algorithm;
  there are no guarantees that it will find the optimum of the posterior pdf in reasonable time.%
  \note{The sampler samples the pdf fairly.
    The tallest peak in the posterior is not guaranteed to be---%
    and in general is not---%
    also the peak with the greatest posterior ``mass''.
    That is, the samples in a fair sampling will not necessarily come
    predominantly from the \emph{tallest} peak in the posterior pdf.
    They will come from the peak with the greatest total integral
    of the likelihood over the prior.
    These issues might sound ``academic'',
    but when the number of parameters gets large (greater than, say, 10),
    many normal intuitions one might have (if any) about what is reasonably likely in a pdf
    are regularly violated.}
For two, the closeness of the best sample to the posterior mode is a strong function of sample size,
  and with very bad scaling.%
  \note{If you have a $K$-point sampling in a $D$-dimensional parameter space,
    there is some best sample $\pars_k^{\best}$.  Now imagine that you want to make
    that best sample 10 times better---that is, you want to find a point $\pars^{\better}$
    that is ten times closer to the true optimum.  How many more samples do you need to take?
    In the limit that $K$ is large, on average, you have to take a factor of
    $10^D$ more samples!
    Any reasonable optimizer is far, far better than this (in terms of compute time).}
For three, if you just want an optimum, use an optimizer!
That is, we \emph{strongly advise against} reporting,
  as ``the'' measurement or ``the result'',
  the parameter value or values for the best sample.
Giving some posterior samples is a good idea (we return to this below).
If you feel drawn to give the best sample,
  instead give the parameter values found by optimizing the posterior PDF,
  starting at the best sample as an initialization
  (and sacrifice your probabilism\note{something}).
This optimal-posterior parameter value is called the
  ``maximum \foreign{a posteriori}'' or ``MAP'' value for the parameters;
  it is like a maximum-likelihood value but regularized by the prior.
It is an estimator with some good (and some bad) properties, but it is not the point of \MCMC,
  and therefore outside the scope of this \documentname.

@@

In principle the only output from the sampling that safely gives both
probabilistic information about the result of the inference and also
good-fitting models is a few---randomly chosen---example samples.
We strongly recommend this\note{HOGG:  And we do this WHERE? Cite Lang, cite eccentricity; etc??};
in particular this is a better thing to do
than to give the ``best sample'', which isn't even guaranteed to be
near the bulk of the posterior pdf or the samples therefrom.

@@

As we have discussed above (\sectionname~\ref{sec:MH}),
\MCMC\ methods are proven to be correct ``in the limit''; that is, when
an infinite number of samples have been taken.
When the function $f(\pars)$ has challenging properties, the results
can approach this limit very slowly (or not at all\note{HOGG: discuss
  situations in which \MCMC\ will not work---related to support.}).
For this reason, in the finite duration of a scientific project, most
users of \MCMC\ have one of two (related) problems:
Either \textsl{(1)}~it is taking too long---it is taking too many steps or
executing too many calls of the function $f(\pars)$ to get independent
samples---or \textsl{(2)}~it is not
exploring the full parameter space---there are local optima (or bad
local geometry around those optima) ``trapping'' the algorithm.

@@

\paragraph{Ensembles:}
As we discussed (\sectionname~\ref{sec:tuning}), one of the critical challenges in the use of
\MCMC\ is tuning the proposal pdf $q(\pars'\given\pars)$.
In problems in which all of the parameters (all of the components or entries of $\pars$) are somehow ``equivalent'',
or the parameters can be seen as components of a vector in a vector
space, there are \emph{ensemble methods} that make use of not just one
random-walker but instead many walkers to automatically generate a properly tuned
proposal distribution.
These methods make use of the distribution of a set of independent
walkers in the parameter space ($\pars$-space) to gauge the typical step-sizes and
directions at which new proposals should be made, obviating or greatly
reducing the tuning requirements.
We are co-authors on one of these methods, \project{emcee}\note{HOGG:
  ref emcee}, which is popular in the astrophysical sciences.
Because ensemble methods are updating independent \MCMC\ chains, they
are often easy to parallelize to make use of multiple cores.

@@

A trivial kind of Gibbs sampling was mentioned above (\sectionname~\ref{sec:MH}) in the
discussion of \MH\ \MCMC:
Even when running vanilla \MH\ \MCMC, it is often useful to update only
one parameter at a time; that is, do only axis-aligned moves in the
$\pars$-space, cycling through axes as you go from step to step.
This helps with tuning and diagnosis, but it also permits a good
implementation to capitalize on simpler calculations for the simpler
parameter directions.\note{%
  The point is that sometimes there are some parameters that only
  affect a very small number of data points, or parameters that change
  the model or the likelihood function and prior values in a very
  simple-to-compute way.  If such simple parameters exist, it makes
  sense to write your likelihood and prior code to capitalize on these
  simplicities.  This is not always trivial, because it may involve
  caching parts of the calculation and so on.  Details would be beyond
  our scope.}

@@

@@

\section{Troubleshooting and advice}\label{sec:trouble}
